### 2024-04-25 chap_RobPol/chap_RobPol.tex
SUBMISSION

\section{Definition of the principal components of the transition}
\label{sec:RobPol:PC_transition}
As detailed in Section \ref{subsec:meth:PCA:transition}, we have decided to define the directions of variation, \ie the robustness metrics, based on the installed capacities through the transition in the different end-use sectors, \ie electricity, \gls{HT} heat, \gls{LT} heat, passenger mobility, freight mobility, \gls{HVC}, ammonia and methanol. These capacities represent the technological roadmaps to supply these \gls{EUD} while respecting the \ce{CO2}-budget.  Data considered in this method come from the \gls{GSA} carried out on the perfect foresight optimisation of the Belgian transition pathway (see Chapter \ref{chap:atom_mol}). This gave 1260 different transitions resulting, for each of them, from the pathway optimisation subject to a sample of uncertain parameters (see Section \ref{subsec:uncert_charac}). Appendix \ref{app:UQ_tech_cap} gives the exhaustive distributions of the installed capacities among the different end-use sectors from the \gls{GSA}.

@@

Finally, we consider the respective contribution of the different technologies in the different $\text{PC}_{y}$, \ie their corresponding component in the different eigenvectors. Highlighting the top-5 technologies for $\text{PC}_{y,1}$, $\text{PC}_{y,2}$ and $\text{PC}_{y,3}$, we observe general trends over the whole transition as well as tipping year where there is a clear trade-off between several technologies (see Figure \ref{fig:Top5_PC_year}). As pointed out in Section \ref{subsec:meth:PCA:transition}, \gls{PCA} does not make any distinction between a vector of variation and its opposite. This is why $\text{PC}_{2025,1}$ and $\text{PC}_{2035,1}$ are actually very similar even though there are mostly on the opposite sides of the 0-axis.

@@

Averaging the components of similar $\text{PC}_y$ allows constructing the $\text{PC}_{\text{transition}}$ (see Figure \ref{fig:Top5_PC_transition}). These directions of variation form the robustness matrix on which it is possible to project the results of transition roadmaps tested under uncertainties and myopic pathway optimisation.

>>>>

\section{Definition of the principal components of the transition}
\label{sec:RobPol:PC_transition}
The directions of variation, \ie the robustness metrics, are based on the installed capacities through the transition in the different end-use sectors, \ie electricity, \gls{HT} heat, \gls{LT} heat, passenger mobility, freight mobility, \gls{HVC}, ammonia and methanol (see Section \ref{subsec:meth:PCA:transition}). These capacities represent the technological roadmaps to supply these \gls{EUD} while respecting the \ce{CO2}-budget.  Data considered in this method come from the \gls{GSA} carried out on the perfect foresight optimisation of the Belgian transition pathway (see Chapter \ref{chap:atom_mol}). This gave 1260 different transitions resulting, for each of them, from the pathway optimisation subject to a sample of uncertain parameters (see Section \ref{subsec:uncert_charac}). Appendix \ref{app:UQ_tech_cap} gives the exhaustive distributions of the installed capacities among the different end-use sectors from the \gls{GSA}.

@@

Finally, we consider the respective contribution of the different technologies in the different $\text{PC}_{y}$, \ie their corresponding component in the different eigenvectors. Highlighting the top-5 technologies for $\text{PC}_{y,1}$, $\text{PC}_{y,2}$ and $\text{PC}_{y,3}$, we observe general trends over the whole transition as well as tipping year where there is a clear trade-off between several technologies (see Figure \ref{fig:Top5_PC_year}). The top-5 technologies of $\text{PC}_{y,1}$ represent on average 99.4\% of the norm of their respective PC. For $\text{PC}_{y,2}$ and $\text{PC}_{y,3}$, this share decreases to 97.6\% and 90.9\%, respectively.  As pointed out in Section \ref{subsec:meth:PCA:transition}, \gls{PCA} does not make any distinction between a vector of variation and its opposite. This is why $\text{PC}_{2025,1}$ and $\text{PC}_{2035,1}$ are actually very similar even though there are mostly on the opposite sides of the 0-axis.

@@

\newpage
Averaging the components of similar $\text{PC}_y$ allows constructing the $\text{PC}_{\text{transition}}$ (see Figure \ref{fig:Top5_PC_transition}). These directions of variation form the robustness matrix on which it is possible to project the results of transition roadmaps tested under uncertainties and myopic pathway optimisation.

### 2024-04-25 chap_atom_mol/chap_atom_mol.tex
RL chapter: analysis about import of electrofuels
SUBMISSION

In their review, \citet{yue2018review} pointed out that uncertainties were accounted for either snapshot models (\ie optimizing a single target future year), to assess a single energy sector (\ie the power sector, most of the time) or with limited number of uncertain parameters, \ie about ten, in a stochastic programming approach to optimize a transition pathway with a small number of time stages, \ie two or three. Here, the uncertainty quantification addresses the total transition pathway of the whole-energy system, with 34 uncertain parameters (see Appendix \ref{app:UC_full}). Accounting for uncertainties presented in Section \ref{sec:meth:UQ}, Section \ref{sec:atom_mol:results_uq} will identify the key factors driving higher or lower imports of electrofuels as well as the installation of \gls{SMR}, by the end of the transition, \ie 2050.

@@

\section{Deterministic impact of integrating \gls{SMR} in 2040}
\label{sec:atom_mol:results_deter} 
In this section, like in the rest of the thesis, the \textbf{REF} case is without any deployment of \gls{SMR} anytime during the transition and whereas in the \textbf{SMR} case this technology is available, up to 6\,GW (see Chapter \ref{chap:case_study}), from 2040 onward. After investigating the deployment of \gls{SMR} through the power sector, the first part of this section focuses on this impact on the whole-system level considerations (\ie overall transition costs, primary energy mix and yearly emissions per sector). The second part will address the impact of \gls{SMR} on each of the other sectors of the system.

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Elec_Tech_Cap_REF.pdf}
\includegraphics[width=0.8\textwidth]{Elec_Tech_Cap_SMR.pdf}
\caption{As soon as available (\ie 2040), \acrfull{SMR} is deployed to their maximum potential (\ie 6\,GW) to substitute more expensive flexible generation units (\ie gas and ammonia \gls{CCGT}).}
\label{fig:results_deter_tech_cap_elec}
\end{figure}

When assessing the electricity production-versus-consumption-balance (\Cref{fig:results_deter_layer_elec}), \gls{SMR}, as a cheaper, flexible and low-emitting power generation system, produces to its full capacity. Given the 15\% maintenance off-time assumed in this work, this represents 44.6\,TWh by 2050. In comparison, in 2020, conventional nuclear power plants produced 34.4\,TWh Belgium. This resurgence of nuclear electricity occurs at the expense of other, although more efficient, technologies: \gls{CCGT} and industrial \gls{CHP}. Besides the unchanged end-use-demand compared with the REF case, we observe a slight increase of the electrification of the rest of the system: +9.4\% which corresponds to +5.8\,TWh, mostly consumed by electric heaters in industry (+48\%) to produce industrial high-temperature heat.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Elec_Layer_REF.pdf}
\includegraphics[width=0.8\textwidth]{Elec_Layer_SMR.pdf}
\caption{The production of electricity from \gls{SMR} substitutes more efficient technologies (\ie \gls{CCGT} and \gls{CHP}) and boosts the electrification of the rest of the system, mostly the industrial high-temperature heat sector.}
\label{fig:results_deter_layer_elec}
\end{figure}

\subsection{System-level impacts}
\label{subsec:atom_mol:results_deter_overall}
First of all, the 6\,GW \gls{SMR} installed from 2040 allow reaching a 3.3\% (-36.9\,b€) cheaper overall transition (\Cref{fig:results_deter_overall_emissions_sector}). Over the 30-year transition, this represents an annual cost saving equal to 0.2\% of the Belgian GDP. Thanks to the perfect foresight approach, the model knows ahead that cheaper and low-emitting \gls{SMR} will be available in the future. As the model can freely spend the constrained \ce{CO2}-budget over the transition, cost-savings also occur before 2040. These early-stage cost-savings are equally shared between the extended use of cheaper \gls{LFO} to produce \gls{HVC} and the delayed deployment of \gls{PV}. Then, the capital-intensive investments in \gls{SMR}, mostly recovered by the end of the transition as salvage value, are widely compensated by the smaller resource-related OPEX. This leads, at the end, aggregating the OPEX and the annualised CAPEX, to a system that is yearly 4.4\,b€ (8.8\%) cheaper by 2050 than in the REF case (Figure \ref{fig:results_deter_overall_emissions_sector}).

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Cum_total_cost_diff_REF.pdf}
\includegraphics[width=0.8\textwidth]{System_cost_DIFF.pdf}
\caption{(Top) Cumulative transition cost and (down) annual system cost differences between the SMR and the REF cases. Including \gls{SMR} ends up in a cheaper overall transition (-36.9\,b€) and cheaper whole-energy system by 2050 (-4.4\,b€).}
\label{fig:results_deter_overall_emissions_sector}
\end{figure}

Extending the use of naphtha-cracker and, to a lesser extent the postponing the deployment of PV, lead to higher \ce{CO2}-emissions at the beginning of the transition that are then compensated by the deployment of \gls{SMR} (\Cref{fig:GWP_per_sector_diff}).

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{GWP_per_sector_diff.pdf}
\caption{Emissions per energy sector differerence between the SMR and the REF cases. The deployment of \gls{SMR} from 2040 onward allows maintaining longer the use of \acrfull{LFO} for the production of \acrfull{HVC}.}
\label{fig:GWP_per_sector_diff}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Primary_mix_REF.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_SMR.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_diff_uranium.pdf}
\caption{(Top) Primary energy mix over the transition for the REF and SMR cases. (Down) Difference of the mix between the two cases after aggregation per category. In line with \citet{rixhon2021terminology},uranium is considered as a non-renewable resource. It is nevertheless dissociated from other non-renewable resources for the sake of clarity. The imported electricity is split between imported non-renewable and renewable resources between the current share of renewables, \ie 37.41\% \cite{eurostat_share_re_elec}, and an assumed 100\%-renewable European electricity mix by 2050.}
\label{fig:results_deter_energy_mix}
\end{figure}

\subsection{Non-power sectors}
\label{subsec:atom_mol:results_deter_others}

@@

Then, as the driving factor for the annualisation and the salvage value of the assets, the interest rate has a 12\% Sobol' index. The model considers an overall interest rate for the entire system (\ie 1.5\% as a nominal value). In practice, the interest rate would vary depending on the technology investment risk. This variation would have, for instance, a major impact on the \gls{LCOE} of technologies like nuclear power plants \cite{world_nuclear_asso}, given the important capital needs and long time horizons \cite{IEA_Nuclear_2022}.

Finally, the cost of purchasing fossil fuels is also a key parameter in terms of variation of the total transition cost. However, due to the ambitious \ce{CO2}-budget, phasing out of fossil fuels is urgent and makes their impact smaller than their renewable alternatives.

@@

@@

Then, the imported e-ammonia rapidly becomes cost-competitive against its fossil alternative (see Chapter \ref{chap:case_study}). At early stage of the transition, e-ammonia substitutes fossil ammonia and the Haber-Bosch process. Where the initial purpose of ammonia is to satisfy a relatively limited \acrfull{NED} (\ie 10 $\pm$ 3\,TWh by 2050), the variation of its import is mostly due to the higher or lower need for ammonia-\gls{CCGT} as a flexible option to produce electricity. From 2035, out of the four considered electrofuels, the imported e-ammonia is the one exhibiting the largest uncertainty with, for instance, an interquartile range (IQR)\footnote{The interquartile range is the difference between the third quartile ($Q3$ or $P_{75}$) and the first ($Q1$ or $P_{25}$). It is an indicator of statistical dispersion around the median, $Q2$ or $P_{50}$.} of about 50\,TWh. In some extreme cases, e-ammonia is the most imported molecules, \ie up to 216\,TWh or 58\% of the total primary mix.

@@

As aforementioned, the industrial \gls{EUD} impacts the most the import of e-methane (Figure \ref{fig:results_uq_samples_gas}). This parameter directly dictates the demand of industrial high-temperature heat for which industrial gas \gls{CHP}, and industrial gas boilers to a lower extent, represent, on average over all the samples, respectively 25.6\% and 6.1\% of the total production. Then, considering the smaller-impact parameters, we notice that \gls{SMR} plays a non-negligible role. Indeed, if deployed, \gls{SMR} produces abundant low-emitting electricity for industrial electric heaters that substitute, even completely in some cases, gas alternatives. This confirms the observation made in Section \ref{subsec:atom_mol:results_deter_others}. Highly available local biomass also leads to smaller import of e-methane to supply bio-hydrolysis and produce methane-equivalent gas. Finally, and surprisingly, costs of purchasing electrofuels and fossil fuels have a positive and negative correlation with the amount of e-methane, respectively. In other words, by 2050, more expensive electrofuels result in more e-methane import. On the contrary, cheaper electrofuels result in higher import of fossil methane. Given the techno-economic optimum sought by EnergyScope, if electrofuels are more expensive, the system will, overall, import less of them, especially e-ammonia, mainly used by \gls{CCGT}. Subject to the \ce{CO2}-budget for the transition, the system goes towards more efficient technologies, like industrial methane-\gls{CHP} to substitute e-ammonia-\gls{CCGT} in the production of electricity. First running on fossil gas, these \gls{CHP} consume more e-methane by 2050. On the contrary, if electrofuels are cheaper, there is more import of them, and especially of e-ammonia. This leaves room for more emitting and cheaper resources to be used while respecting the \ce{CO2}-budget, \ie coal in industrial boilers that produce, in these cases, on average 24\% of the high-temperature (HT) heat in 2050. In these cases, the use of coal in 2050 highlights that, with a sharper decrease of the emissions at early stages in the transition, the model finds a solution including highly-emitting resources (\eg coal) while respecting the \ce{CO2}-budget. Consequently, there is smaller investment in methane \gls{CHP}, and consequently import of e-methane as more abundant renewable electricity is produced via e-ammonia-\gls{CCGT} and more HT-heat is supplied by industrial coal boilers. Even though we might expect that no more coal will be consumed in Belgium by 2050, the model still has the opportunity to use it if the \ce{CO2}-budget allows it. About the cost of purchasing fossil fuels, the parameter has mainly an impact on the import of fossil \gls{NG} as the most versatile energy carrier in the whole-energy system. If \gls{NG} is more expensive, the system will import less of it. Subsequently, the investments in methane-\gls{CHP} and boilers are more limited. This ends up in smaller need for e-methane by 2050.

@@

@@

\subsection{Local renewables}
\label{subsec:atom_mol:results_uq_VRES}
In line with results given in Section \ref{sec:atom_mol:results_deter}, the \gls{GSA} shows that \gls{SMR} has negligible impact on the deployment of local \gls{VRES} (\ie \gls{PV}, onshore and offshore wind turbines). Whereas the installed capacities of onshore wind is totally driven by the uncertainty on its maximum potential, Sobol' indices change for the most impacting parameters on the deployment of \gls{PV} and offshore wind (see Figure \ref{fig:results_uq_pdf_local_ren}). The key factor that drives the installed capacities of these two technologies is mostly their respective maximum potential, especially at the end of the transition, much more than their CAPEX. Given its higher \gls{LCOE} (\Cref{fig:LCOE}), \gls{PV} is more impacted in the short-term by the variation on the cost of purchasing electrofuels supplying e-methane (and e-ammonia to a lesser extent) \gls{CCGT}. However, this impact gets negligible by 2050.

>>>>

\section*{Contributions}
\label{sec:atom_mol:contributions}

In their review, \citet{yue2018review} pointed out that uncertainties were accounted for either snapshot models (\ie optimizing a single target future year), to assess a single energy sector (\ie the power sector, most of the time) or with limited number of uncertain parameters, \ie about ten, in a stochastic programming approach to optimise a transition pathway with a small number of time stages, \ie two or three. Here, the uncertainty quantification addresses the total transition pathway of the whole-energy system, with 34 uncertain parameters (see Section \ref{sec:cs:uncertainty}). Accounting for uncertainties presented in Section \ref{sec:meth:UQ}, Section \ref{sec:atom_mol:results_uq} will identify the key factors driving higher or lower imports of electrofuels as well as the installation of \gls{SMR}, by the end of the transition, \ie 2050.

@@

\section[Deterministic impact of integrating SMR in 2040]{Deterministic impact of integrating \gls{SMR} in 2040}
\label{sec:atom_mol:results_deter} 
In this section, like in the rest of the thesis, the \textbf{REF} case is without any deployment of \gls{SMR} anytime during the transition and whereas in the \textbf{SMR} case this technology is available, up to 6\,GW (see Chapter \ref{chap:case_study}), from 2040 onward. After investigating the deployment of \gls{SMR} through the power sector, the first part of this section focuses on this impact on the whole-system level considerations (\ie overall transition costs, primary energy mix and yearly emissions per sector). The second part will address the impact of \gls{SMR} on each of the other sectors of the system.

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Elec_Tech_Cap_REF.pdf}
\includegraphics[width=\textwidth]{Elec_Tech_Cap_SMR.pdf}
\caption{As soon as available (\ie 2040), \acrfull{SMR} is deployed to their maximum potential (\ie 6\,GW) to substitute more expensive flexible generation units (\ie gas and ammonia \gls{CCGT}).}
\label{fig:results_deter_tech_cap_elec}
\end{figure}

\newpage
When assessing the electricity production-versus-consumption-balance (\Cref{fig:results_deter_layer_elec}), \gls{SMR}, as a cheaper, flexible and low-emitting power generation system, produces to its full capacity. Given the 15\% maintenance off-time assumed in this work, this represents 44.6\,TWh by 2050. In comparison, in 2020, conventional nuclear power plants produced 34.4\,TWh Belgium. This resurgence of nuclear electricity occurs at the expense of other, although more efficient, technologies: \gls{CCGT} and industrial \gls{CHP}. Besides the unchanged end-use-demand compared with the REF case, we observe a slight increase of the electrification of the rest of the system: +9.4\% which corresponds to +5.8\,TWh, mostly consumed by electric heaters in industry (+48\%) to produce industrial high-temperature heat.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Elec_Layer_REF.pdf}
\includegraphics[width=\textwidth]{Elec_Layer_SMR.pdf}
\caption{The production of electricity from \gls{SMR} substitutes more efficient technologies (\ie \gls{CCGT} and \gls{CHP}) and boosts the electrification of the rest of the system, mostly the industrial high-temperature heat sector.}
\label{fig:results_deter_layer_elec}
\end{figure}

\newpage
\subsection{System-level impacts}
\label{subsec:atom_mol:results_deter_overall}
First of all, the 6\,GW \gls{SMR} installed from 2040 allow reaching a 3.3\% (-36.9\,b€) cheaper overall transition (\Cref{fig:results_deter_overall_emissions_sector}). Over the 30-year transition, this represents an annual cost saving equal to 0.2\% of the Belgian GDP. Thanks to the perfect foresight approach, the model knows ahead that cheaper and low-emitting \gls{SMR} will be available in the future. As the model can freely spend the constrained \ce{CO2}-budget over the transition, cost-savings also occur before 2040. These early-stage cost-savings are equally shared between the extended use of cheaper \gls{LFO} to produce \gls{HVC} and the delayed deployment of \gls{PV}. Then, the capital-intensive investments in \gls{SMR}, mostly recovered by the end of the transition as salvage value, are widely compensated by the smaller resource-related OPEX. This leads, at the end, aggregating the OPEX and the annualised CAPEX, to a system that is yearly 4.4\,b€ (8.8\%) cheaper by 2050 than in the REF case (Figure \ref{fig:results_deter_overall_emissions_sector}).

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Cum_total_cost_diff_REF.pdf}
\includegraphics[width=\textwidth]{System_cost_DIFF.pdf}
\caption{(Top) Cumulative transition cost and (down) annual system cost differences between the SMR and the REF cases. Including \gls{SMR} ends up in a cheaper overall transition (-36.9\,b€) and cheaper whole-energy system by 2050 (-4.4\,b€).}
\label{fig:results_deter_overall_emissions_sector}
\end{figure}

Extending the use of naphtha-cracker and, to a lesser extent, postponing the deployment of PV, lead to higher \ce{CO2}-emissions at the beginning of the transition that are then compensated by the deployment of \gls{SMR} (\Cref{fig:GWP_per_sector_diff}).

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{GWP_per_sector_diff.pdf}
\caption{Emissions per energy sector differerence between the SMR and the REF cases. The deployment of \gls{SMR} from 2040 onward allows maintaining longer the use of \acrfull{LFO} for the production of \acrfull{HVC}.}
\label{fig:GWP_per_sector_diff}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Primary_mix_REF.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_SMR.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_diff_uranium.pdf}
\caption{(Top) Primary energy mix over the transition for the REF and SMR cases. (Bottom) Difference of the mix between the two cases after aggregation per category. In line with \citet{rixhon2021terminology},uranium is considered as a non-renewable resource. It is nevertheless dissociated from other non-renewable resources for the sake of clarity. The imported electricity is split between imported non-renewable and renewable resources between the current share of renewables, \ie 37.41\% \cite{eurostat_share_re_elec}, and an assumed 100\%-renewable European electricity mix by 2050.}
\label{fig:results_deter_energy_mix}
\end{figure}

\newpage
\subsection{Non-power sectors}
\label{subsec:atom_mol:results_deter_others}

@@

@@

As the driving factor for the annualisation and the salvage value of the assets, the interest rate has a 12\% Sobol' index. The model considers an overall interest rate for the entire system (\ie 1.5\% as a nominal value). In practice, the interest rate would vary depending on the technology investment risk. This variation would have, for instance, a major impact on the \gls{LCOE} of technologies like nuclear power plants \cite{world_nuclear_asso}, given the important capital needs and long time horizons \cite{IEA_Nuclear_2022}.

The cost of purchasing fossil fuels is also a key parameter in terms of variation of the total transition cost. However, due to the ambitious \ce{CO2}-budget, phasing out of fossil fuels is urgent and makes their impact smaller than their renewable alternatives.

@@

Then, the imported e-ammonia rapidly becomes cost-competitive against its fossil alternative (see Chapter \ref{chap:case_study}). At early stage of the transition, e-ammonia substitutes fossil ammonia and the Haber-Bosch process. Where the initial purpose of ammonia is to satisfy a relatively limited \acrfull{NED} (\ie 10 $\pm$ 3\,TWh by 2050), the variation of its import is mostly due to the higher or lower need for ammonia-\gls{CCGT} as a flexible option to produce electricity. From 2035, out of the four considered electrofuels, the imported e-ammonia is the one exhibiting the largest uncertainty with, for instance, an interquartile range (IQR)\footnote{The interquartile range is the difference between the third quartile ($Q3$ or $P_{75}$) and the first ($Q1$ or $P_{25}$). It is an indicator of statistical dispersion around the median, $Q2$ or $P_{50}$.} of about 50\,TWh. In some extreme cases, e-ammonia is the most imported molecules, \ie up to 167\,TWh or 45\% of the total primary mix.

@@

@@

As aforementioned, the industrial \gls{EUD} impacts the most the import of e-methane (Figure \ref{fig:results_uq_samples_gas}). This parameter directly dictates the demand of industrial high-temperature heat for which industrial gas \gls{CHP}, and industrial gas boilers to a lower extent, represent, on average over all the samples, respectively 25.6\% and 6.1\% of the total production. Then, considering the smaller-impact parameters, we notice that \gls{SMR} plays a non-negligible role. Indeed, if deployed, \gls{SMR} produces abundant low-emitting electricity for industrial electric heaters that substitute, even completely in some cases, gas alternatives. This confirms the observation made in Section \ref{subsec:atom_mol:results_deter_others}. Highly available local biomass also leads to smaller import of e-methane to supply bio-hydrolysis and produce methane-equivalent gas. Finally, and surprisingly, costs of purchasing electrofuels and fossil fuels have a positive and negative correlation with the amount of e-methane, respectively. In other words, by 2050, more expensive electrofuels result in more e-methane import. On the contrary, cheaper electrofuels result in higher import of fossil methane. Given the techno-economic optimum sought by EnergyScope, if electrofuels are more expensive, the system will, overall, import less of them, especially e-ammonia, mainly used by \gls{CCGT}. Subject to the \ce{CO2}-budget for the transition, the system goes towards more efficient technologies, like industrial methane-\gls{CHP} to substitute e-ammonia-\gls{CCGT} in the production of electricity. First running on fossil gas, these \gls{CHP} consume more e-methane by 2050. On the contrary, if electrofuels are cheaper, there is more import of them, and especially of e-ammonia. This leaves room for more emitting and cheaper resources to be used while respecting the \ce{CO2}-budget, \ie coal in industrial boilers that produce, in these cases, on average 24\% of the high-temperature (HT) heat in 2050. In these cases, the use of coal in 2050 highlights that, with a sharper decrease of the emissions at early stages in the transition, the model finds a solution including highly-emitting resources (\eg coal) while respecting the \ce{CO2}-budget. Consequently, there are smaller investments in methane \gls{CHP}, and consequently import of e-methane as more abundant renewable electricity is produced via e-ammonia-\gls{CCGT} and more HT-heat is supplied by industrial coal boilers. Even though we might expect that no more coal will be consumed in Belgium by 2050, the model still has the opportunity to use it if the \ce{CO2}-budget allows it. About the cost of purchasing fossil fuels, the parameter has mainly an impact on the import of fossil \gls{NG} as the most versatile energy carrier in the whole-energy system. If \gls{NG} is more expensive, the system will import less of it. Subsequently, the investments in methane-\gls{CHP} and boilers are more limited. This ends up in smaller need for e-methane by 2050.

@@

\newpage
\subsection{Local renewables}
\label{subsec:atom_mol:results_uq_VRES}
In line with results given in Section \ref{sec:atom_mol:results_deter}, the \gls{GSA} shows that \gls{SMR} has negligible impact on the deployment of local \gls{VRES} (\ie \gls{PV}, onshore and offshore wind turbines). Whereas the installed capacities of onshore wind is totally driven by the uncertainty on its maximum potential, Sobol' indices change for the most impacting parameters on the deployment of \gls{PV} and offshore wind (see Figure \ref{fig:results_uq_pdf_local_ren}). The key factor that drives the installed capacities of these two technologies is mostly their respective maximum potential, especially at the end of the transition, much more than their CAPEX. Given its higher \gls{LCOE} (\Cref{fig:LCOE}), \gls{PV} is more impacted in the short-term by the variation on the cost of purchasing electrofuels supplying e-methane (and e-ammonia to a lesser extent) \gls{CCGT}. However, this impact gets negligible by 2050.
